\MCnote{Reachability-Based Safety and Goal Satisfaction of Unmanned Aerial Vehicle Platoons on Air Highways}

### 2016-11-09 docs/AIAA 2016 highway platooning/conclusions.tex
edited up to reviewer 2: air highways

% !TEX root = main.tex
\section{Conclusions}
To address the important and urgent problem of the traffic management of unmanned aerial vehicles (UAVs), we proposed to have platoons of UAVs traveling on air highways. We showed how such an airspace structure leads to much easier safety and liveness analysis. We provided simulations which show that by organizing vehicles into platoons, many complex maneuvers can be performed using just a few different reachable sets.

>>>>

% !TEX root = main.tex
\section{Conclusions}
To address the important and urgent problem of the traffic management of unmanned aerial vehicles (UAVs), we proposed to have platoons of UAVs traveling on air highways. We showed how such an airspace structure leads to much easier safety and liveness analysis. We provided simulations which show that by organizing vehicles into platoons, many complex maneuvers can be performed using just a few different backward reachable sets.

### 2016-11-09 docs/AIAA 2016 highway platooning/high_form.tex
edited up to reviewer 2: air highways

\MCnote{Note that if vehicles in the airspace are far away from each other, it may be reasonable for all vehicles to remain in the ``free'' mode and fly in an unstructured manner. As long as multiple-way conflicts do not occur, pairwise collision avoidance maneuvers would be sufficient to ensure safety. Unstructured flight is likely to result in more efficient trajectories for each individual vehicle.}

>>>>

\MCnote{In addition, if vehicles in the airspace are far away from each other, it may be reasonable for all vehicles to remain in the ``free'' mode and fly in an unstructured manner. As long as multiple-way conflicts do not occur, pairwise collision avoidance maneuvers would be sufficient to ensure safety. Unstructured flight is likely to result in more efficient trajectories for each individual vehicle.}

### 2016-11-09 docs/AIAA 2016 highway platooning/introduction.tex
edited up to reviewer 2: air highways

>>>>

\MCnote{Overall, this paper is not meant to provide an exhaustive solution to the airspace management problem. Instead, this paper illustrates that the computation intractability of HJ reachability can be overcome using an air highway structure with UAVs flying in platoons. In addition, the results are intuitive, which can facilitate human participation in managing the airspace. Although many challenges not addressed in this paper still need to be overcome, this paper can provide a starting point for future research in large-scale UASs with safety and goal satisfaction guarantees.}

### 2016-11-09 docs/AIAA 2016 highway platooning/plat_safety.tex
edited up to reviewer 2: air highways

% !TEX root = main.tex
\subsection{Safety Analysis}
Under normal operations in a single platoon, each follower vehicle $\veh{i},i=P_2,\ldots,P_{M-1}$ in a platoon checks whether it is in the safety reachable set with respect to $\veh{P_{i-1}}$ and $\veh{P_{i+1}}$. So $\vehSCS{i} = \{P_{i+1}, P_{i-1}\}$ for $i=P_2,\ldots,P_{N-1}$. Assuming there are no nearby vehicles outside of the platoon, the platoon leader $\veh{P_1}$ checks safety against $\veh{P_2}$, and the platoon trailer $Q_{P_N}$ checks safety against $Q_{P_{N-1}}$. So $\vehSCS{P_1}=\{P_2\}, \vehSCS{P_N}=\{P_{N-1}\}$. When all vehicles are using goal satisfaction controllers to perform their allowed maneuvers, in most situations no pair of vehicles should be in an unsafe configuration. However, occasionally a vehicle $\veh{k}$ may behave unexpectedly due to faults or malfunctions, in which case it may come into an unsafe configuration with another vehicle.

With our choice of $\vehSCS{i}$ and the assumption that the platoon is in a single-file formation, some vehicle $\veh{i}$ would get near the safety reachable set with respect to $\veh{k}$, where $\veh{k}$ is likely to be the vehicle in front or behind of $\veh{i}$. In this case, a ``safety breach" occurs. Our synthesis of the safety controller guarantees that between every pair of vehicles $\veh{i},\veh{k}$, as long as $V_S(-t,x_i- x_k)>0$, $\exists u_i$ to keep $\veh{i}$ from colliding with $\veh{k}$ for a desired time horizon $t$, despite the worst case (an adversarial) control from $\veh{k}$. Therefore, as long as the number of ``safety breaches" is at most one for $\veh{i}$, $Q_i$ can simply use the optimal control to avoid $\veh{k}$ and avoid collision for the time horizon of $t$. Under the assumption that vehicles are able to exit the current altitude range within a duration of $\td$, if we choose $t=\td$, the safety breach would always end before any collision can occur.

>>>>

% !TEX root = main.tex
\subsection{Safety Analysis}
Under normal operations in a single platoon, each follower vehicle $\veh{i},i=P_2,\ldots,P_{M-1}$ in a platoon checks whether it is in the safety BRS with respect to $\veh{P_{i-1}}$ and $\veh{P_{i+1}}$. So $\vehSCS{i} = \{P_{i+1}, P_{i-1}\}$ for $i=P_2,\ldots,P_{N-1}$. Assuming there are no nearby vehicles outside of the platoon, the platoon leader $\veh{P_1}$ checks safety against $\veh{P_2}$, and the platoon trailer $Q_{P_N}$ checks safety against $Q_{P_{N-1}}$. So $\vehSCS{P_1}=\{P_2\}, \vehSCS{P_N}=\{P_{N-1}\}$. When all vehicles are using goal satisfaction controllers to perform their allowed maneuvers, in most situations no pair of vehicles should be in an unsafe configuration. However, occasionally a vehicle $\veh{k}$ may behave unexpectedly due to faults or malfunctions, in which case it may come into an unsafe configuration with another vehicle.

With our choice of $\vehSCS{i}$ and the assumption that the platoon is in a single-file formation, some vehicle $\veh{i}$ would get near the safety BRS with respect to $\veh{k}$, where $\veh{k}$ is likely to be the vehicle in front or behind of $\veh{i}$. In this case, a ``safety breach" occurs. Our synthesis of the safety controller guarantees that between every pair of vehicles $\veh{i},\veh{k}$, as long as $V_S(-t,x_i- x_k)>0$, $\exists u_i$ to keep $\veh{i}$ from colliding with $\veh{k}$ for a desired time horizon $t$, despite the worst case (an adversarial) control from $\veh{k}$. Therefore, as long as the number of ``safety breaches" is at most one for $\veh{i}$, $Q_i$ can simply use the optimal control to avoid $\veh{k}$ and avoid collision for the time horizon of $t$. Under the assumption that vehicles are able to exit the current altitude range within a duration of $\td$, if we choose $t=\td$, the safety breach would always end before any collision can occur.

### 2016-11-09 docs/AIAA 2016 highway platooning/plat_sim.tex
edited up to reviewer 2: air highways

The first two plots in Figure \ref{fig:fp} illustrate the use of goal satisfaction and safety reachable sets for the first two vehicles. Since the goal satisfaction reachable sets are in 4D and the safety reachable sets are in 6D, we compute and plot their 2D slices based on the vehicles' velocities and relative velocities.  All vehicles begin as Free vehicles, so they each need to take into account five different reachable sets: four safety reachable sets and one goal satisfaction reachable set. For clarity, we only show the goal satisfaction reachable set and the four safety reachable sets for one of the vehicles.

@@

At $t=4.2$, $\veh{1}$ (orange) is inside the goal satisfaction reachable set for getting to an absolute state, shown as the dotted orange boundary. Therefore, it is ``locked-in" to the target state $\bar{x}_H$, and follows the optimal control in \eqref{eq:HJB_ctrl_syn} to $\bar{x}_H$. During the entire time, $\veh{1}$ checks whether it may collide with any of the other vehicles within a time horizon of $\td$. To do this, it simply checks whether its state relative to each of the other vehicles is within the corresponding safety reachable set. As an example, the safety reachable set boundary with respect to $\veh{2}$ (purple) is shown as the orange dashed boundary around $\veh{2}$ (purple); $\veh{1}$ (orange) is safe with respect to $\veh{2}$ (purple) since $\veh{1}$ (orange) is outside of the boundary. In fact, $\veh{1}$ is safe with respect to all vehicles.

@@

At $t=8.0$, $Q_2$ (purple) is in the process of joining the platoon behind $\veh{1}$ (orange) by moving towards the target $\bar{x}_P$ relative to the position of $\veh{1}$ (orange). Note that $\bar{x}_P$ moves with $\veh{1}$ (orange) as $\bar{x}_P$ is defined in terms of the relative states of the two vehicles. Since $\veh{2}$ is inside the goal satisfaction reachable set boundary for joining the platoon (purple dotted boundary), it is ``locked-in" to the target relative state $\bar{x}_P$, and begins following the optimal control in \eqref{eq:HJI_ctrl_syn} towards the target as long as it stays out of all safety reachable sets. For example, at $t=5.9$, $\veh{2}$ (purple) is outside of the safety reachable set with respect to $\veh{1}$ (orange), shown as the purple dashed boundary around $\veh{1}$ (orange). Again, from the other safety reachable set boundaries, we can see that $\veh{2}$ is in fact safe with respect to all vehicles.

@@

Figure \ref{fig:in} shows the simulation result. At $t=9.9$, a platoon of four vehicles, $\veh{i},i=1,\ldots,4$ (with $P_i = i$), travels along the highway shown. An intruder vehicle $\veh{5}$ (yellow) heads left, disregarding the presence of the platoon. At $t=11.9$, the platoon leader $\veh{1}$ (red) detects that it has gone near the boundary of the safety reachable set (not shown) with respect to the intruder $\veh{5}$ (yellow). In response, $\veh{1}$ (red) starts using the safety controller to optimally avoid the intruder according to \eqref{eq:HJI_ctrl_syn}; in doing so, it steers slightly off the highway.

Note that although in this particular simulation, the intruder travels in a straight line, a straight line motion of the intruder was \textit{not} assumed. Rather, the safety reachable sets are computed assuming the worst case control of the intruder, according to \eqref{eq:HJI_ctrl_syn}.

As the intruder $\veh{5}$ (yellow) continues to disregard other vehicles, the followers of the platoon also get near the respective boundaries of their safety reachable sets with respect to the intruder. This occurs at $t=13.9$, where the platoon ``makes room'' for the intruder to pass by to avoid collisions; all vehicles deviate from their intended path, which is to follow the platoon leader or the highway.

After the intruder has passed, eventually all vehicles become far away from any safety reachable sets. When this occurs, the leader resumes following the highway, and the followers resume following the leader. At $t=19.9$, the platoon successfully gets back onto the highway.

% 100 120 140 200
\begin{figure}
    \centering
    \begin{subfigure}[t]{0.45\columnwidth} \label{subfig:in_100}
        \includegraphics[width=\columnwidth]{fig/in_100}
        \caption{A four-vehicle platoon travels along the highway. The yellow vehicle disregards the others.}
    \end{subfigure}
    \begin{subfigure}[t]{0.45\columnwidth} \label{subfig:in_120}
        \includegraphics[width=\columnwidth]{fig/in_120}
        \caption{Vehicles begin avoidance maneuvers as they get near safety reachable set boundaries.}
    \end{subfigure}

>>>>

The first two plots in Figure \ref{fig:fp} illustrate the use of safety and goal satisfaction BRS for the first two vehicles. Since the goal satisfaction BRSs are in 4D and the safety BRSs are in 6D, we compute and plot their 2D slices based on the vehicles' velocities and relative velocities.  All vehicles begin as Free vehicles, so they each need to take into account five different BRSs: four safety BRSs and one goal satisfaction BRS. For clarity, we only show the goal satisfaction BRS and the four safety BRSs for one of the vehicles.

@@

At $t=4.2$, $\veh{1}$ (orange) is inside the goal satisfaction BRS for getting to an absolute state, shown as the dotted orange boundary. Therefore, it is ``locked-in" to the target state $\bar{x}_H$, and follows the optimal control in \eqref{eq:HJB_ctrl_syn} to $\bar{x}_H$. During the entire time, $\veh{1}$ checks whether it may collide with any of the other vehicles within a time horizon of $\td$. To do this, it simply checks whether its state relative to each of the other vehicles is within the corresponding safety BRS. As an example, the safety BRS boundary with respect to $\veh{2}$ (purple) is shown as the orange dashed boundary around $\veh{2}$ (purple); $\veh{1}$ (orange) is safe with respect to $\veh{2}$ (purple) since $\veh{1}$ (orange) is outside of the boundary. In fact, $\veh{1}$ is safe with respect to all vehicles.

@@

At $t=8.0$, $Q_2$ (purple) is in the process of joining the platoon behind $\veh{1}$ (orange) by moving towards the target $\bar{x}_P$ relative to the position of $\veh{1}$ (orange). Note that $\bar{x}_P$ moves with $\veh{1}$ (orange) as $\bar{x}_P$ is defined in terms of the relative states of the two vehicles. Since $\veh{2}$ is inside the goal satisfaction BRS boundary for joining the platoon (purple dotted boundary), it is ``locked-in" to the target relative state $\bar{x}_P$, and begins following the optimal control in \eqref{eq:HJI_ctrl_syn} towards the target as long as it stays out of all safety BRSs. For example, at $t=5.9$, $\veh{2}$ (purple) is outside of the safety BRS with respect to $\veh{1}$ (orange), shown as the purple dashed boundary around $\veh{1}$ (orange). Again, from the other safety BRS boundaries, we can see that $\veh{2}$ is in fact safe with respect to all vehicles.

@@

Figure \ref{fig:in} shows the simulation result. At $t=9.9$, a platoon of four vehicles, $\veh{i},i=1,\ldots,4$ (with $P_i = i$), travels along the highway shown. An intruder vehicle $\veh{5}$ (yellow) heads left, disregarding the presence of the platoon. At $t=11.9$, the platoon leader $\veh{1}$ (red) detects that it has gone near the boundary of the safety BRS (not shown) with respect to the intruder $\veh{5}$ (yellow). In response, $\veh{1}$ (red) starts using the safety controller to optimally avoid the intruder according to \eqref{eq:HJI_ctrl_syn}; in doing so, it steers slightly off the highway.

Note that although in this particular simulation, the intruder travels in a straight line, a straight line motion of the intruder was \textit{not} assumed. Rather, the safety BRSs are computed assuming the worst case control of the intruder, according to \eqref{eq:HJI_ctrl_syn}.

As the intruder $\veh{5}$ (yellow) continues to disregard other vehicles, the followers of the platoon also get near the respective boundaries of their safety BRSs with respect to the intruder. This occurs at $t=13.9$, where the platoon ``makes room'' for the intruder to pass by to avoid collisions; all vehicles deviate from their intended path, which is to follow the platoon leader or the highway.

After the intruder has passed, eventually all vehicles become far away from any safety BRSs. When this occurs, the leader resumes following the highway, and the followers resume following the leader. At $t=19.9$, the platoon successfully gets back onto the highway.

% 100 120 140 200
\begin{figure}
    \centering
    \begin{subfigure}[t]{0.45\columnwidth} \label{subfig:in_100}
        \includegraphics[width=\columnwidth]{fig/in_100}
        \caption{A four-vehicle platoon travels along the highway. The yellow vehicle disregards the others.}
    \end{subfigure}
    \begin{subfigure}[t]{0.45\columnwidth} \label{subfig:in_120}
        \includegraphics[width=\columnwidth]{fig/in_120}
        \caption{Vehicles begin avoidance maneuvers as they get near safety BRS boundaries.}
    \end{subfigure}
